---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# sospatial

The goal of sospatial is to provide some easy to use Southern Ocean data sets for 
mapping and exploration in R. 

We aim to fill some existing gaps where (especially) mapping in polar regions causes
difficulties. 

## Example

This example shows the used of the `packed_lats` cache to draw a rough map of the intensity
of where the daily sea ice extent was in the Southern Ocean. 

WIP - when this data set is fully spatially and temporally referenced we can use it 
to extract abstract metrics like "max sea ice extent", which might mean the major 'edge' on the day of greatest extent, or the cumulatively most-northerly location at all longitudes during a year ...


```{r example}
## reconstruct record
library(dplyr)
library(sospatial)
ice <- tibble(lon = rep_len(seq(-180, 179), length.out = length(packed_lats)), 
              lat = packed_lats/10,
              day = rep(ice_dates, each = 360))

ice[c("x", "y")] <- rgdal::project(as.matrix(ice[c("lon", "lat")]), "+proj=stere +lat_0=-90 +datum=WGS84")
library(ggplot2)
ggplot(ice, aes(x, y)) + geom_bin2d() + coord_equal()

# what does a month look like
mon <- ice %>% dplyr::filter(format(day, "%m") == "08")
ggplot(mon, aes(x, y)) + geom_bin2d() + coord_equal() + xlim(range(mon$x)) + ylim(range(mon$y))

mon$era <- c("old", "new")[(mon$day > as.POSIXct("1998-06-15")) + 1]
ggplot(mon, aes(x, y)) + 
  geom_bin2d() + coord_equal() + xlim(range(mon$x)) + ylim(range(mon$y)) + 
  facet_wrap(~era)

```


# Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

